---
title: "Functions"
output: html_document
---
###Function Definition

Functions are written like below (usually in their own file). The value of the last statement gets returned.

```{r}
myAdd <- function(x, y){
  x + y
}

myAdd(2,3)
```

Functions are first class objects and functions can be nested.

Parameters can also have default values.

```{r}
above <- function(x, n = 10){
  x[x>n]
}

above(1:20,3)
above(1:20)
```

###Argument Matching
Arguments can be matched by position or by name (or by a mixture of that)
```{r}
myAdd <- function(x, y){
  x + y
}

myAdd(2,3)
myAdd(y = 3, x = 2)
```

That is usefull for functions with long argument lists where one does not want to specify all parameters
```{r}
args(lm)
```

Arguments can also be partially matched
```{r}
myAdd <- function(myLongParameterNamme, y){
  myLongParameterNamme + y
}

myAdd(y = 3, m = 2)
```

The matching order is:

* Exact match by name
* Partial match by name
* Match by position

###Argument Evaluation

Arguments are evaluated lazily. So the following does not lead to an error
```{r}
myFun <- function(x, y){
  x^2
}

myFun(3)
```

### The "..." Argument

... indicates a variable number of arguments. Usually that is used to pass them to another function.
```{r}
pasteWithDashes <- function(...){
  paste(..., sep="-")
}

pasteWithDashes('a', 'b', 'c')
```

Everything after ... argument must be named explicitely.