---
title: "Functions"
output: html_document
---
###Function Definition

Functions are written like below (usually in their own file). The value of the last statement gets returned.

```{r}
myAdd <- function(x, y){
  x + y
}

myAdd(2,3)
```

Functions are first class objects and functions can be nested.

Parameters can also have default values.

```{r}
above <- function(x, n = 10){
  x[x>n]
}

above(1:20,3)
above(1:20)
```

###Argument Matching
Arguments can be matched by position or by name (or by a mixture of that)
```{r}
myAdd <- function(x, y){
  x + y
}

myAdd(2,3)
myAdd(y = 3, x = 2)
```

That is usefull for functions with long argument lists where one does not want to specify all parameters
```{r}
args(lm)
```

Arguments can also be partially matched
```{r}
myAdd <- function(myLongParameterNamme, y){
  myLongParameterNamme + y
}

myAdd(y = 3, m = 2)
```

The matching order is:

* Exact match by name
* Partial match by name
* Match by position

###Argument Evaluation

Arguments are evaluated lazily. So the following does not lead to an error
```{r}
myFun <- function(x, y){
  x^2
}

myFun(3)
```

### The "..." Argument

... indicates a variable number of arguments. Usually that is used to pass them to another function.
```{r}
pasteWithDashes <- function(...){
  paste(..., sep="-")
}

pasteWithDashes('a', 'b', 'c')
```

Everything after ... argument must be named explicitely.

### Environments
#####From the R Language Definition
Environments can be thought of as consisting of two things: a frame, which is a set of symbol-value pairs, and an enclosure, a pointer to an enclosing environment. When R looks up the value for a symbol the frame is examined and if a matching symbol is found its value will be returned. If not, the enclosing environment is then accessed and the process repeated. Environments form a tree structure in which the enclosures play the role of parents. The tree of environments is rooted in an empty environment, available through emptyenv(), which has no parent.

So environments are a place to store variables. If ones starts up r and makes an assignment a variable in the global environment:
```{r}
rm(list = ls())
a <- 1
ls()
ls(globalenv())
```



###Functions and Environments (Lexical Scoping)
#####From the R Language Definition
Functions (or more precisely, function closures) have three basic components: a formal argument list, a body and an environment. [...] A function's environment is the environment that was active at the time that the function was created. [...] When a function is called, a new environment (called the evaluation environment) is created, whose enclosure is the environment from the function closure. This new environment is initially populated with the unevaluated arguments to the function; as evaluation proceeds, local variables are created within it.

Lets create a function within a function and see what environments are created:

```{r}
rm(list = ls())

f <- function(){
  g <- function(){
  }
}
f()
environment(f)
```

The invornment of f is the global environenment:
```{r}
environment(f)
```

The environment of g however is the environment created when f was called:
```{r}
g <- f()
environment(g)
```
The parent of g's environment is f's environment which is the global environment again:
```{r}
parent.env(environment(g))
```

####Lexical Scoping
That now gets interesting together with free variables in functions:
```{r}
rm(list = ls())
a <- 2000
b <- 1000
x <- 1
f <- function(){
  a <- 200
  b <- b <- 100
  y <- 2
  g <- function(){
    a <- 20
    x + y + a + b
  }
}
g <- f()
```
The environments do now contain some variables
```{r}
ls(environment(g))
ls(environment(f))
```
Note that a and b are contained in different environments but with different values
```{r}
environment(g)$a
environment(f)$a
environment(g)$b
environment(f)$b
```
If we now call g r tries to assign values to the variables for the calculation (x + y + a + b). Therefore it looks through the environments starting with the environment that is created when g is called and if it doesnt find a symbol, it moves up to the parent.
```{r}
g()
```
If it reaches the empty environment, it throws an exception
```{r error=TRUE}
rm(x)
g()
```