---
title: "Typesystem"
output: html_document
---


###Vectors
The basic unit of data in R is the vector. (Even a single integer is represented as a vector of length 1)

R has six basic (‘atomic’) vector types:

* logical
* integer
* real
* complex
* string (or character)
* raw

```{r}
str(c(TRUE, FALSE))
str(c(1L, 2L))
str(c(1, 1.1))
str(c(1i, 1+1i))
str(c('a', 'b'))
```

###Vector manipulation
####Creation
```{r}
c(1,2,3,5,8) #generate a vector with specific elements
1:10 # generate a sequence-vector
vector("numeric", length=10)  # generate a numerical vector of length 10 with default values
```
###Functions that help us to explore R objects
```{r}
vec <- c(1,2,3)
str(vec) # structure of an object 
length(vec) # number of elements or components
class(vec)  # class or type of an object
summary(vec) # different information about the object
attributes(matrix(1:2)) # print attributes of the object
table(c("a","a","b")) # interpret input data as factors and do a frequency count
```
####Automatic coercion
If you mix the datatypes in the vector R is going to convert them to the least common datatype
```{r}
str(c(1,"a")) # 1 is converted to chr
str(c(TRUE, 1)) # TRUE is converted to num
```
###Numbers
If you just type a number it is always treated as num (real number)
```{r}
str(1)
```
For explicitly havin an integer, you need the suffix L
```{r}
str(1L)
```
A special number is Inf, which means infinity
```{r}
1/0
```
and there is also NaN, which means not a number.
```{r}
0/0
```
###Data conversations
You can convert objects to different types with the as.* functions
```{r}
vec <- c(1,2,3)
str(vec)
as.numeric(vec)
as.character(vec)
```
###Matrices
Matrices are vector with a dimension attribute.
```{r}
mat <- matrix(nrow = 3, ncol = 4) # create a matix of dimension 3x4
mat
attributes(mat) # The dim attribute is itself a vector of lenght 2
dim(mat) # Get the dimension of the matrix
mat <- matrix(1:12, nrow = 3, ncol = 4) # Matrices are constructed by filling it column by column
mat
cbind(1:3, 6:8) # create a matrix by binding two column vectors together
rbind(1:3, 6:8) # create a matrix by binding two row vectors together
```
###Lists
Special types of vectors that can contain different types
```{r}
l <- list(1, "a", TRUE, 1+1i)
l
```
###Factors
Factors are used to represent categorical data. They can be ordered or unordered.
```{r}
facts <- factor(c("big","medium","small","big", "small"))
facts
table(facts) # Get a frequency count of the labels
facts <- factor(c("big","medium","small","big", "small"),
                levels = c("small", "medium", "big")) # construct factors with specific label ordering
facts
```
###Missing Values
Missing values are denoted by NA or NaN
```{r}
is.na(NA) # function to test for NA of an object
is.nan(NaN) # function to test for NaN of an object
is.na(NaN) # NaN are also NA
is.nan(NA) # but NA are not NaN
is.na(c(1, NA, 2)) # produce a logical vector that indicates wether an element is NA or not
```
###Data Frames
Data frames are used to store tabular data.
```{r}
df <- data.frame(nr = 1:3, name = c("John", "Joe", "Jamie")) # construct a simple dataframe
df
nrow(df) # number of rows
ncol(df) # number of columns
```
